Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning

### 2013-10-30 04_experiment.tex
rephrased intro, amended layout
Sections 2,3 proof

In order to minimize the effect of the confounding factors, we counterbalanced the positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions. Thus the unique color in a context could be either blue (as in Fig. \ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e. a context always has one square and two circles. }.

@@

There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}. First of all, the original experiment used the betting paradigm, i.e. participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct. Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action. This can lead to confusion when we are to fit the model predictions to the empirical data without knowing whether they are directly comparable. In addition, since the betting paradigm is more or less introspective in nature, it tends to be not very accurate. Thus we used forced choice in our experiment instead, which clearly measures the action and does not rely on introspection.
Secondly, since we decided to investigate the influence of the speaker's preference as well as the listener's perceptual salience, we focused on contexts equivalent to Figure \ref{context}, which are most typical in Gricean accounts of scalar implicature, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure the vocabulary to be common knowledge, excluding the feature of texture. Finally, we did not use dotted line surrounding an object as the way to indicate the target to the speaker, as it could emphasize the feature of shape and thus add confound in the experiment. Instead we used a small arrow pointing to the target.

@@

>>>>

In order to minimize the effect of the confounding factors, we counterbalanced the positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions. Thus the unique color in a context could be either blue (as in Fig. \ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e., a context always has one square and two circles. }.

@@

There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}. First of all, the original experiment used the betting paradigm, i.e., participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct. Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action. This can lead to confusion when we are to fit the model predictions to the empirical data without knowing whether they are directly comparable. In addition, since the betting paradigm is more or less introspective in nature, it tends to be not very accurate. Thus we used forced choice in our experiment instead, which clearly measures the action and does not rely on introspection.
Secondly, since we decided to investigate the influence of the speaker's preference as well as the listener's perceptual salience, we focused on contexts equivalent to Figure \ref{context}, which are most typical in Gricean accounts of scalar implicature, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure the vocabulary to be common knowledge, excluding the feature of texture. Finally, we did not use dotted line surrounding an object as the way to indicate the target to the speaker, as it could emphasize the feature of shape and thus add confound in the experiment. Instead we used a small arrow pointing to the target.

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End:

### 2013-10-30 06_conclusion.tex
rephrased intro, amended layout
Sections 2,3 proof

Our experiment and model comparison suggest the following: (1) On the conceptual level, 
it is helpful to clarify the distinction between belief, goal and action, and the Bayesian framework provides us with a natural way to structure these components into a formal model. By examining each component systematically, we can explicitly spell out the underlying assumptions and formulate alternative hypotheses to be tested. (2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience. Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions. There is some correlation between the speaker's and the listener's rationality,  while either side has his own bias and is unaware of the other's. Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting. Understanding the relation between the forced-choice and the betting paradigms would be an important next step.

>>>>

Our experiment and model comparison suggest the following: (1) On the
conceptual level, it is helpful to clarify the distinction between
belief, goal and action, and the Bayesian framework provides us with a
natural way to structure these components into a formal model. By
examining each component systematically, we can explicitly spell out
the underlying assumptions and formulate alternative hypotheses to be
tested. (2) On the empirical level, we tested the RSA model with a
family of variants motivated from the game theoretic pragmatic
perspectives on communication, and in particular, we investigated the
speaker's preference as well as the listener's perceptual
salience. Our data showed a more intricate picture about the roles
various factors play in pragmatic reasoning about referential
expressions. There is some correlation between the speaker's and the
listener's rationality, while either side has his own bias and is
unaware of the other's. Also, an action-based view might better
reflect the goal of communication, at least in the forced-choice
experimental setting. Understanding the relation between the
forced-choice and the betting paradigms would be an important next
step.
